---
title: "1.4 — Data Wrangling in the tidyverse — R Practice"
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
```

- [<i class="fas fa-cloud"></i> View Project on R Studio Cloud](https://rstudio.cloud/spaces/83147/project/1555051)
- [<i class="fas fa-external-link-square-alt"></i> `Answer Key (html)`](/r/1.4-r-practice-answers.html)
- [<i class="fab fa-markdown"></i> `Answer Key (markdown)`](/r/1.4-r-practice-answers.Rmd)
- [<i class="fas fa-registered"></i> `Answer Key (R Script)`](/r/1.4-r-practice-answers.R)

# Getting Set Up

Before we begin, start a new file with `File` $\rightarrow$ `New File` $\rightarrow$ `R Script`. As you work through this sheet in the console in `R`, also add (copy/paste) your commands that work into this new file. At the end, save it, and run to execute all of your commands at once.

First things first, load `tidyverse`.

```{r}
library("tidyverse")
```

# Warm Up to `dplyr` With `gapminder`

## 1.
Load `gapminder`. First, `select()` the variables `year`, `lifeExp`, `country`.

## 2.
`select()` all variables *except* `pop`.

## 3.
`rename()` `continent` to `cont`.

## 4. 
`arrange()` by year.

## 5. 
`arrange()` by year, but in descending order.

## 6. 
`arrange()` by year, then by life expectancy.

## 7. 
`filter()` observations with `pop` greater than 1 billion.

## 8. 
Of those, look only at `India`. 

## 9.
Try out the pipe (`%>%`) if you haven't already, by chaining commands: `select()` your data to look only at `year`, `gdpPercap`, and `country` in the year `1997`, for countries that have a `gdpPercap` greater than 20,000, and `arrange()` them alphabetically.

## 10.
`mutate()` a new variable called `GDP` that is equal to `gdpPercap * pop`.

## 11.
`mutate()` a new population variable that is the `pop` in millions. 

## 12.
`summarize()` to get the average GDP per capita.

## 13.
Get the number of observations, average, minimum, maximum, and standard deviation for GDP per capita.

## 14.
Get the average GDP per capita over time. Hint, first `group_by()` year.

## 15.
Get the average GDP per capita by continent.

## 16.
Get the average GDP per capita by year and by continent.^[Hint: do `year` first, if you do `continent` first, there are no years to group by!] Then save this as another `tibble` called `gdp`. Create a `ggplot` of a `line` graph of average continent GDP over time using the `gdp` data. 

## 17.
Try it again all in one command with the pipe `%>%`. Instead of saving the data as `gdp`, pipe it right into `ggplot`!^[Hint: You can use `.` as a placeholder.] 

# Example: the Economics of College Majors

Now let's step it up to work with some data "in the wild" to answer some research questions. This will have you combine your `dplyr` skills and add some new things such as importing with `readr`. 

Let's look at fivethirtyeight's article ["
The Economic Guide To Picking A College Major
"](https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/). fivethirtyeight is great about making the data behind their articles public, we can [download all of their data here](https://data.fivethirtyeight.com/). Search for `college majors` and click download (the blue arrow button).^[This will download a `.zip` file that contains many spreadsheets. Unzip it with a program that unzips files (such as WinZip, 7-zip, the Unarchiver, etc).] We will look at the `recent-grads.csv` file. 

The description in the `readme` file for the data is as follows:

Header | Description
---|---------
`Rank` | Rank by median earnings
`Major_code` | Major code, FO1DP in ACS PUMS
`Major` | Major description
`Major_category` | Category of major from Carnevale et al
`Total` | Total number of people with major
`Sample_size` | Sample size (unweighted) of full-time, year-round ONLY (used for earnings)
`Men` | Male graduates
`Women` | Female graduates
`ShareWomen` | Women as share of total
`Employed` | Number employed (ESR == 1 or 2)
`Full_time` | Employed 35 hours or more
`Part_time` | Employed less than 35 hours
`Full_time_year_round` | Employed at least 50 weeks (WKW == 1) and at least 35 hours (WKHP >= 35)
`Unemployed` | Number unemployed (ESR == 3)
`Unemployment_rate` | Unemployed / (Unemployed + Employed)
`Median` | Median earnings of full-time, year-round workers
`P25th` | 25th percentile of earnigns
`P75th` | 75th percentile of earnings
`College_jobs` | Number with job requiring a college degree
`Non_college_jobs` | Number with job not requiring a college degree
`Low_wage_jobs` | Number in low-wage service jobs

## 18. 
Import the data with `read_csv()` and assign it to an object called `majors`.^[One way to avoid error messages is to move (on your computer) `recent_grads.csv` to the same folder as R's working directory, which again you can check with `getwd()`.] The first argument of this command is the name of the original file, in quotes.^[If the file is in a different folder, the argument is the full path in quotes.]

## 19.
Look at the data with `glimpse()`. This is a suped-up version of `str()` in `tidyverse`. 

## 20.
What are all of the *unique* values of `Major`? How many are there?

## 21.
Which major has the *lowest* unemployment rate?

## 22. 
What are the top 3 majors that have the highest percentage of women?

### 23. 
Make a boxplot of `Median` wage by `Major_Category`.^[You won't be able to read the labels easily, so add `theme(axis.text.x=element_text(angle=45, hjust=1)` to angle `x`-axis labels (and move them down by 1)]

### 24.
Which major category is the least popular in this sample?^[Hint: use `group_by` first.]


### 25. 
Is there a systematic difference in median earnings between STEM majors and non-STEM majors? First define: 

```{r}
stem_categories <- c("Biology & Life Science",
                     "Computers & Mathematics",
                     "Engineering",
                     "Physical Sciences")
```
Next, make a variable called `stem`, for whether or not a `Major_category` is `"stem"` or `"not_stem"`.^[Hint: try out the `ifelse()` function which has three inputs: condition(s) for a variable(s), what to do if `TRUE` (the if), and what to if `FALSE` (the else), i.e.

```{r, eval=F}
stem = ifelse(my_conditions,
       yes = do_this_if_TRUE,
       no = do_this_if_FALSE)
```

You'll of course need to change the `do_this` into something!
]

Then `summarize()` `median` for stem and not stem groups.