---
title: "3.2 — Causal Inference II: DAGs — R Practice"
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
```

# Set Up
To minimize confusion, I suggest creating a new `R Project` (e.g. `regression_practice`) and storing any data in that folder on your computer.

Alternatively, I have made a project in R Studio Cloud that you can use (and not worry about trading room computer limitations), with the data already inside (you will still need to assign it to an object).

- [<i class="fas fa-external-link-square-alt"></i> Answer Key (html)](/r/3.2-r-practice-answers.html)
- [<i class="fab fa-markdown"></i> Answer Key (markdown)](/r/3.2-r-practice-answers.Rmd)

# DAG Practice

For each of the following examples:

- i. Write out *all* of the causal pathways from `X` (treatment of interest) to `Y` (outcome of interest).
- ii. Identify which variable(s) need to be controlled for to estimate the causal effect of `X` on `Y`. You can use [`dagitty.net`](http://dagitty.net) to help you, but you should start trying to recognize these on your own!
- iii. Draw the DAGs in `r` using `ggdag`. After setting up the dag with `dagify()` (and specifying `exposure` and `outcome` inside `dagify`), pipe that into `ggdag()`. Try again piping it instead into `ggdag_status()` (to highlight what is X and what is Y). Try again piping it instead into `ggdag_adjustment_set()` to show what needs to be controlled.

Don’t forget to install `ggdag`!

```{r, echo=F, message=F, warning=F}
# load to draw DAGs! 
library(tidyverse)
library(ggdag)
```

## Question 1

```{r, echo=F, fig.retina=3}
dagify(Y~X+Z,
       X~Z,
       exposure = "X",
       outcome = "Y") %>% 
  ggdag(stylized = FALSE, seed=1)+theme_dag_blank()+theme(legend.position = "none")
```

## Question 2

```{r, echo=F, fig.retina=3}
dagify(Y~X+M,
       M~X,
       exposure = "X",
       outcome = "Y") %>% 
  ggdag(stylized = FALSE, seed=1)+theme_dag_blank()+theme(legend.position = "none")
```

## Question 3

```{r, echo=F, fig.retina=3}
dagify(Y~X+Z+B,
       Z~B+A,
       X~A,
       exposure = "X",
       outcome = "Y") %>% 
  ggdag(stylized = FALSE, seed=1)+theme_dag_blank()+theme(legend.position = "none")
```

## Question 4

```{r, echo=F, fig.retina=3}
dagify(Y~X+B+C,
       Z~B+A,
       X~A,
       C~X,
       exposure = "X",
       outcome = "Y") %>% 
  ggdag(stylized = FALSE, seed=2)+theme_dag_blank()+theme(legend.position = "none")
```

## Question 5

```{r, echo=F, fig.retina=3}
dagify(Y~X+Z+A,
       Z~X+A+B,
       B~A,
       exposure = "X",
       outcome = "Y") %>% 
  ggdag(stylized = FALSE, seed=2)+theme_dag_blank()+theme(legend.position = "none")
```