---
title: "2.7 — Inference for Regression — R Practice"
author: "Solutions"
date: "ECON 480 — Econometrics — Fall 2020"
output:
  html_document:
    df_print: paged
    theme: simplex
    toc: true 
    toc_depth: 3
    toc_float: true
    code_folding: show
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(20) # using this number means all "random" generated objects will be identical for all of us!
```

## Question 1
Let’s use the `diamonds` data built into `ggplot`. Simply load `tidyverse` and then to be clear, save this as a tibble (feel free to rename it) with `diamonds <- diamonds`.

---

<!--ANSWER BELOW HERE-->

```{r}
library(tidyverse)
diamonds <- diamonds
```

---

## Question 2
Suppose we want to estimate the following relationship:

$$\text{price}_i = \beta_0 + \beta_1 \text{carat}_i + u_i$$

Run a regression of `price` on `carat` using `lm()` and get a `summary`.

---

<!--ANSWER BELOW HERE-->

```{r}
reg <- lm(price ~ carat, data = diamonds)

summary(reg)
```

```{r}
# it will also help to visualize

ggplot(data = diamonds)+
  aes(x = carat,
      y = price)+
  geom_point()+
  geom_smooth(method = "lm")
```

---

### Part A

What is $\hat{\beta_1}$? Interpret it in the context of our regression.

---

<!--ANSWER BELOW HERE-->

$\hat{\beta_1}$, the estimated slope, is 7756.43. For every 1 carat, we expect the price of the diamond to increase by $7,756.43.

---

### Part B
Use `broom`'s `tidy()` command, and calculate a confidence interval by including `conf.int = T` inside `tidy()`. What is the 95% confidence interval for $\hat{\beta_1}$, and what does it mean? Save these endpoints as an object.

---

<!--ANSWER BELOW HERE-->

```{r}
library(broom)

reg_tidy<-reg %>% tidy(conf.int = T)

# look at tidy outcome
reg_tidy

# extract & save the confidence interval
CI_endpoints <- reg_tidy %>%
  filter(term == "carat") %>%
  select(conf.low, conf.high)

# look at it
CI_endpoints
```

We are 95% confident that in similarly constructed samples, the true effect of carat on price $(\beta_1)$ is between $7,728.86 and $7,784.00.


I am also going to extract and save our $\hat{\beta_1}$ estimate as an object, that we will use later with `infer`:

```{r}
# extract and save sample slope as object called beta_1_hat

beta_1_hat <- reg_tidy %>%
  filter(term == "carat") %>%
  select(estimate)

# confirm
beta_1_hat
```

---

## Question 3
Now let’s use `infer`. Install it if you don’t have it, then load it.

---

<!--ANSWER BELOW HERE-->

```{r}
# install.packages("infer")
library(infer)
```

---

### Part A

Let’s generate a confidence interval. First `specify()` the model relationship, then `generate()` `reps = 1000` repetitions of the sample using a `type = bootstrap`, then have it `calculate(stat = "slope")`.^[Note this will take a few minutes, its doing a lot of calculations!] What does it show you?

---

<!--ANSWER BELOW HERE-->

```{r}
diamonds %>%
  specify(price ~ carat) %>%
  generate(reps = 1000,
           type = "bootstrap") %>%
  calculate(stat = "slope")
```

This produces 1,000 bootstrapped samples, and calculates the estimated regression slope $(\hat{\beta_1})$ for each sample.

---

### Part B

Continue the pipeline from part A, next have it `get_confidence_interval()`. Set `level = 0.95, type = "se"` and `point_estimate` equal to our estimated $\hat{\beta_1}$ from Question 2.

---

<!--ANSWER BELOW HERE-->

```{r}
diamonds %>%
  specify(price ~ carat) %>%
  generate(reps = 1000,
           type = "bootstrap") %>%
  calculate(stat = "slope") %>%
  get_confidence_interval(level = 0.95, #<<
                          type = "se", #<<
                          point_estimate = beta_1_hat) #<<

# in the last part, you could type in our actual beta 1 hat, 7756.426;
# but I had saved it for easy use
```

This gives us a confidence interval of $7,706.95 and $7,805.91. This not exactly, but quite close to `R`'s automatic calculation using the theoretical $t$-distribution in `lm()`.

---

### Part C
Now instead of `get_confidence_interval()`, pipe into `visualize()` to see the distribution. If you saved the confidence interval endpoints from part 1B, you can finally add `+shade_ci(endpoints = ...)` setting the argument equal to whatever you called your object containing the confidence interval.

---

<!--ANSWER BELOW HERE-->

```{r}
diamonds %>%
  specify(price ~ carat) %>%
  generate(reps = 1000,
           type = "bootstrap") %>%
  calculate(stat = "slope") %>%
  visualize()+
  shade_ci(endpoints = CI_endpoints)
```

---

## Question 4

Now let’s test the following hypothesis:

$$\begin{align*}
H_0: \beta_1 &= 0\\
H_1: \beta_1 &\neq 0\\
\end{align*}$$

### Part A

What does the output of `summary` or of `tidy` from Question 2 tell you?

---

<!--ANSWER BELOW HERE-->

Reading the second row of estimates from the `lm summary` or `tidy` output of question 2, we saw that the `estimate` on `carat` $(\hat{\beta_1})$ was 7756.43, with a standard error of 14.07. This yields a test statistic value of 551.4, and a $p$-value of basically 0.

If you are wondering, the test-statistic was calculated, as described in the slides, as:

$$\begin{align*}
t & = \frac{\text{estimate} - \text{null hypothesis}}{\text{SE(estimate)}}\\
t & = \frac{7756.43 - 0}{14.07}\\
t &\approx 551.3 \\
\end{align*}$$

If the null hypothesis were true $(\beta_1=0)$, the probability that we get a test-statistic at least as extreme as 551.3 (essentially, 551 standard deviations away!!) is virtually 0. You can see this when we visualize using infer, below.

---

### Part B
Let’s now do this with `infer`. First `specify()` the model relationship, then `hypothesize(null = "independence")` to declare $H_0: \beta_1 = 0$, then `generate()` `reps = 1000` repetitions of the sample using a `type = permute`, then have it `calculate(stat = "slope")`. What does it show you?

---

<!--ANSWER BELOW HERE-->

```{r}
diamonds %>%
  specify(price ~ carat) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000,
           type = "permute") %>%
  calculate(stat = "slope")
```

This produces 1,000 permuted samples, under the assumption that the true slope is 0 (the null hypothesis), and calculates the estimated regression slope $(\hat{\beta_1})$ for each sample.

---

### Part C
Continue the pipeline from part B, next have it `get_p_value()`. Inside this function, set `obs_stat` equal to our $\hat{\beta_1}$ we found, and set `direction = "both"` to run a two-sided test, since our alternative hypothesis is two-sided, $H_1: \beta_1 \neq 0$.

---

<!--ANSWER BELOW HERE-->

```{r}
diamonds %>%
  specify(price ~ carat) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000,
           type = "permute") %>%
  calculate(stat = "slope") %>%
  get_p_value(obs_stat = beta_1_hat, #<<
              direction = "both") #<<

# again, instead of beta_1_hat you could type in 7756.426
```

---

### Part D

Instead of `get_p_value()`, pipe into `visualize(obs_stat = ... , direction = "both").` where `...` is our estimated $\hat{\beta_1}$.

---

<!--ANSWER BELOW HERE-->

```{r}
diamonds %>%
  specify(price ~ carat) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000,
           type = "permute") %>%
  calculate(stat = "slope") %>%
  visualize(obs_stat = beta_1_hat, #<<
                direction = "both") #<<

```
