---
title: "3.9: Logarithmic Models - R Practice"
author: "YOUR NAME"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    #theme: 
    toc: true 
    toc_depth: 3
    toc_float: true
    code_folding: show
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

We are returning to the speeding tickets data that we began to explore in [R Practice 3.4 on Multivariate Regression](http://metricsf20.classes.ryansafner.com/practice/3.4-practice) and [R Practice 3.7 on Dummy Variables & Interaction Effects](http://metricsf20.classes.ryansafner.com/practice/3.7-r-practice). Download and read in (`read_csv`) the data below.

- [<i class="fas fa-table"></i> `speeding_tickets.csv`](https://metricsf20.classes.ryansafner.com/data/speeding_tickets.csv)

This data again comes from a paper by Makowsky and Strattman (2009) that we will examine later. Even though state law sets a formula for tickets based on how fast a person was driving, police officers in practice often deviate from that formula. This dataset includes information on all traffic stops. An amount for the fine is given only for observations in which the police officer decided to assess a fine. There are a number of variables in this dataset, but the one's we'll look at are:

| Variable | Description |
|----------|-------------|
| `Amount` | Amount of fine (in dollars) assessed for speeding |
| `Age` | Age of speeding driver (in years) |
| `MPHover` | Miles per hour over the speed limit |
| `Black` | Dummy $=1$ if driver was black, $=0$ if not |
| `Hispanic` | Dummy $=1$ if driver was Hispanic, $=0$ if not |
| `Female` | Dummy $=1$ if driver was female, $=0$ if not |
| `OutTown` | Dummy $=1$ if driver was not from local town, $=0$ if not |
| `OutState` | Dummy $=1$ if driver was not from local state, $=0$ if not |
| `StatePol` | Dummy $=1$ if driver was stopped by State Police, $=0$ if stopped by other (local) |

> We again want to explore who gets fines, and how much.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 2

Run a regression of `Amount` on `Age`. Write out the estimated regression equation, and interpret the coefficient on Age.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 3

Is the effect of `Age` on `Amount` nonlinear? Let's run a quadratic regression. 

### Part A

Create a new variable for $Age^2$. Then run a quadratic regression.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part B

Try running the same regression using the alternate notation: `lm(Y~X+I(X^2))`. This method allows you to not have to create a new variable first. Do you get the same results?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part C

Write out the estimated regression equation.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part D

Is this model an improvement from the linear model?^[Check $R^2$.]

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part E

Write an equation for the marginal effect of `Age` on `Amount`.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part F

Predict the marginal effect on `Amount` of being one year older when you are 18.  How about when you are 40?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part G

Our quadratic function is a $U$-shape. According to the model, at what age is the amount of the fine minimized?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part H

Create a scatterplot between `Amount` and `Age` and add a a layer of a linear regression (as always), and an additional layer of your predicted quadratic regression curve. The regression curve, just like any regression *line*, is a `geom_smooth()` layer on top of the `geom_point()` layer. We will need to customize `geom_smooth()` to `geom_smooth(method="lm", formula="y~x+I(x^2)` (copy/paste this verbatim)! This is the same as a regression line (`method="lm"`), but we are modifying the formula to a polynomial of degree 2 (quadratic): $y=a+bx+cx^2$.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part I

It's quite hard to see the quadratic curve with all those data points. Redo another plot and this time, only keep the quadratic `stat_smooth()` layer and leave out the `geom_point()` layer. This will only plot the regression curve.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 4

Should we use a higher-order polynomial equation? Run a cubic regression, and determine whether it is necessary.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 5

Run an $F$-test to check if a nonlinear model is appropriate. Your null hypothesis is $H_0: \beta_2=\beta_3=0$ from the regression in pert (h). The command is `linearHypothesis(reg_name, c("var1", "var2"))` where `reg_name` is the name of the `lm` object you saved your regression in, and `var1` and `var2` (or more) in quotes are the names of the variables you are testing. This function requires (installing and) loading the "`car`" package (additional regression tools).

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 6

Now let's take a look at speed (`MPHover` the speed limit).

### Part A
Creating new variables as necessary, run a **linear-log** model of `Amount` on `MPHover`. Write down the estimated regression equation, and interpret the coefficient on `MPHover` $(\hat{\beta_1})$. Make a scatterplot with the regression line.^[Hint: The simple `geom_smooth(method="lm")` is sufficient, so long as you use the right variables on the plot!]

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part B

Creating new variables as necessary, run a **log-linear** model of `Amount` on `MPHover`. Write down the estimated regression equation, and interpret the coefficient on `MPHover` $(\hat{\beta_1})$. Make a scatterplot with the regression line.^[Hint: The simple `geom_smooth(method="lm")` is sufficient, so long as you use the right variables on the plot!]

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part C

Creating new variables as necessary, run a **log-log** model of `Amount` on `MPHover`. Write down the estimated regression equation, and interpret the coefficient on `MPHover` $(\hat{\beta_1})$. Make a scatterplot with the regression line.^[Hint: The simple `geom_smooth(method="lm")` is sufficient, so long as you use the right variables on the plot!]

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part D

Which of the three log models has the best fit?^[Hint: Check $R^2$]

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 7

Return to the quadratic model. Run a quadratic regression of `Amount` on `Age`, `Age`$^2$, `MPHover`, and all of the race dummy variables. Test the null hypothesis: *"the race of the driver has no effect on Amount"*

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 8

Now let's try standardizing variables. Let's try running a regression of `Amount` on `Age` and `MPHover`, but standardizing each variable. 

### Part A

Create new standardized variables for `Amount`, `Age`, and `MPHover`.^[Hint: use the `scale()` function inside of `mutate()`]

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part B

Run a regression of standardized `Amount` on standardized `Age` and `MPHover`. Interpret $\hat{\beta_1}$ and $\hat{\beta_2}$ Which variable has a bigger effect on `Amount`? 
