---
title: "1.3 — Data Visualization with ggplot2 — Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Tuesday, August 25, 2020*

# Overview

Today we start the fun stuff - data visualization. We will cover how to build plots with the package `ggplot2` (part of the `tidyverse`): I will lecture for the first half (again, it will be a lot of information that you can refer back to as needed over the semester) and you will practice making plots.

# Readings

Please see [today's readings](/reading/1.3-reading) for more information and helpful resources about all of your data visualization questions. The last slide today also contains a list of helpful online resources.

# Slides

- [<i class="fas fa-chalkboard-teacher"></i> View Lecture Slides](/slides/1.3-slides.html)
- [<i class="fas fa-file-pdf"></i> Download as PDF](/slides/1.3-slides.pdf)

# Live Class Session on Zoom

The live class <i class="fas fa-video"></i> Zoom meeting link can be found on Blackboard (see `LIVE ZOOM MEETINGS` on the left navigation menu), starting at 11:30 AM.

If you are unable to join today's live session, or if you want to review, you can find the recording stored on Blackboard via Panopto (see `Class Recordings` on the left navigation menu).

# Practice Problems

Answers from last class’ practice problems on base R are posted [on that page](/r/1.2-r-practice). Today you will be working on [R practice problems](/r/1.3-r-practice) on data visualization. Answers will be posted later.

# Statistics and Software Survey

[“Answers”](/assignment/preliminary-survey-answers.html) for the statistics survey have been posted, including some summaries of how your classmates feel.

# Assignments

None!

# Additional Useful Information

## New Packages Mentioned

Learn more about each of these in context in the slides, but below are links to each packages’s online help pages, which include descriptions of how each function works, examples, and cheatsheets to help you

- [`ggplot2`](https://ggplot2.tidyverse.org/) a grammar of graphics for ploting

Check back here later for more tips on plotting with ggplot2!

## Extensions to ggplot2

`ggplot2`, being one of the most popular packages, has a lot of user-made extensions that allow you to do lots of neat things with your plots, from plotting networks, Parliaments, dendrograms, and other types of graphs, to formatting and visual tools that help improve your figures.

For the following demonstrations, we will use the `gapminder` data. Let’s start just by making a basic graph and saving it as an object called `p`. I have decided to map (`aes()`) each `geom_point`’s `color` to `continent` and `size` to `pop`:

```{r}
library(gapminder)
library(ggplot2)

p<-ggplot(gapminder) +
  aes(x = gdpPercap,
      y = lifeExp) +
  geom_point(aes(size = pop,
                 color = continent))
p
```

### Working with Scales

I don't like the default choices `ggplot2` has made for my point sizes for population, or the way it depicts them (in scientific notation) on the legend. 

I will set my own scale by setting the `breaks`^[The cut offs for using one size circle vs. the next/previous size circle, depending on population.] manually, according to a vector I define as: `c(100000, 1000000, 100000000, 1000000000)`. So, I will use one point size for populations of 100 thousand, a bigger one for a million, a bigger one for 100 millions, and the biggest for 1 billion. 

I am going to `label` these (on my legend) as the following vector: `c("<1 million","1 million","100 million", "1 billion")`.

Lastly, I don’t think the size of the billion circle is big enough, a billion is a lot of people! So I will set the `range` of sizes from size `1` point to size `10` point.

To do this, I include all of this inside the `scale_size` command (because I am scaling the `size` of points):

```{r}
# let's save this as p2
p2<-p+scale_size(breaks = c(100000, 1000000, 100000000, 1000000000), # cut offs
             labels=c("<1 million","1 million","100 million", "1 billion"), # labels on legend
             range=c(1,10)) # min & max point size

# let's see what we did
p2
```

This is also very hard to see the relationship (because it is nonlinear!). So I will rescale the `x_axis` logarithmically with base 10:

```{r}
p2+scale_x_log10()
```

Doing this already gives me a much clearer view of the relationship! But I don't like the labels, or the breaks it has chosen, so I can customize them again:

```{r}
p2+scale_x_log10(
    breaks = c(10^3, 10^4, 10^5), # 1,000, 10,000, and 100,000
    labels = scales::dollar)
```

The `scales` package has a nice command to format labels, in this case I am calling the `scales::dollar` function to print dollar signs in front of my axes numbers. I could have done it manually instead by setting `labels = c("$1,000", "$10,000", "$100,000")`.

## Subsetting Data

We learn more about this in [class 1.4](/class/1.4-class) using `tidyverse`, but let’s only look at one year of data (there’s too much going on in this plot, especially with the large points, some points are covering other points). So let’s only look at the year 2007. I can do this in two ways:

1. Subset the data, save the subsetted data as an object (I'll call `gap2007`), plot with that object as my data:

```{r}
library(tidyverse)

gap2007 <- gapminder %>%
  filter(year == 2007)

p2007 <- ggplot(data = gap2007)+
  aes(x = gdpPercap,
      y = lifeExp) +
  geom_point(aes(size = pop,
                 color = continent))

p2007
```

2. Subset data and pipe it directly into `ggplot2`’s `data` argument:

```{r}
library(tidyverse)

p2007 <- gapminder %>%
  filter(year == 2007) %>%
  ggplot(data = .)+ # . is placeholder
  aes(x = gdpPercap,
      y = lifeExp) +
  geom_point(aes(size = pop,
                 color = continent))

p2007
```

Now let’s clean up the graph with the same things I did before, hide the legends (set the `color` and `size`, my two `aes`thetic mappings, `guides` equal to `FALSE`), add some labels, and change the theme:

```{r}
p3<-p2007+scale_size(breaks = c(100000, 1000000, 100000000, 1000000000), 
             labels=c("<1 million","1 million","100 million", "1 billion"), 
             range=c(1,10))+
  scale_x_log10(
    breaks = c(10^3, 10^4, 10^5),
    labels = scales::dollar)+
  labs(x = "GDP per Capita (USD)",
       y = "Life Expectancy (years)")+
  guides(color = FALSE,
         size = FALSE)+
  theme_classic()
p3
```

## ggrepel

If I were to try to label some countries, with either `geom_text` (just word) or `geom_label()` (text in a box), setting the `label` `aes`thetic to `country`, see what would happen:

```{r}
p3+geom_label(aes(label = country,
                  color = continent))
```

The labels, which are plotted right on top of each point, cover the points!

Someone figured out a clever way to let us do both, and it is a package called `ggrepel`, which allows you to plot labels that are "repelled" away from the point they are labelling in an intelligent way. This is a separate package, which you must first *install* and then load with *library* to use it!

```{r}
# install.packages("ggrepel") # do this only once
library(ggrepel)

p3+geom_text_repel(aes(label = country,
                        color = continent),
                    size = 3)
```

This is much better, but for this particular chart, since a lot of observations are close together, it would be unwise to label *everything*, perhaps only label a subset of important points.

## ggflag

One alternative is instead of points, to use some other marking. Someone created the `ggflags` package to let you plot flags of countries. This creates a new type of geom, called `geom_flag`, that requires you to map the `country` `aes`thetic to a variable in your data with the country name (incidentally, in `gapminder` that variable is also called `country`). Let’s try that out instead (and add my same customizations as above):

```{r}
# install.packages("ggflags") # do this only once
library(ggflags)
pflag<-p3+geom_flag(aes(country = country))
pflag
```

## plotly

We can also make our plot a bit more interactive (on web only of course!) using the `ggplotly` package, which allows `ggplot2` to interface with a javascript library called `plotly`.

```{r}
# install.packages("ggplotly") # do this only once
library(plotly)
ggplotly(pflag)
```