---
title: "2.2 — Random Variables and Distributions — Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Tuesday, September 8, 2019*

# Overview

Today we finish your crash course/review of basic statistics with random variables and distributions.

# Slides

- [<i class="fas fa-chalkboard-teacher"></i> View Lecture Slides](/slides/2.1-slides.html)
- [<i class="fas fa-file-pdf"></i> Download as PDF](/slides/2.1-slides.pdf)

# Live Class Session on Zoom

The live class <i class="fas fa-video"></i> Zoom meeting link can be found on Blackboard (see `LIVE ZOOM MEETINGS` on the left navigation menu), starting at 11:30 AM.

If you are unable to join today's live session, or if you want to review, you can find the recording stored on Blackboard via Panopto (see `Class Recordings` on the left navigation menu).

# Problem Set

[Problem Set 1](/assignment/01-problem-set) answers are posted on that page.

[Problem set 2](/assignment/02-problem-set) (on classes 2.1-2.2) is posted shortly, and is will be due by Sunday September 13.

# Appendix: Graphing Statistical and Mathematical Functions in R

The `mosaic` package is useful for making and using mathematical functions in `R`.

```{r, message = F, warning = F}
# install.packages("mosaic") install for first use
library(mosaic) # load 
```

## Creating Mathematical Functions

You can create custom mathematical functions using mosaic by defining an R `function()` with multiple arguments. As a simple example, make the function $f(x) = 10x-x^2$ as follows:

```{r}
# store as a named function, I'll call it "my_function"
my_function<-function(x){10*x-x^2}

# look at it
my_function

```

There are some notational requirements from `R` for making functions. Any coefficient in front of a variable (such as the 10 in `10x` must be explicitly multiplied by the variable, as in `10*x`).

To use the function, simply define what the input `(x)` is and run your named function on it: 

```{r}
# f of 2 
my_function(2)

# f of 2 and 4
my_function(c(2,4))

# f of 2 through 7
my_function(2:7)

# ALTERNATIVELY
# define x first and then run function on it

x<-c(2,4)
my_function(x)
```

## Graphing Mathematical Functions

In `ggplot` there is a dedicated `stat_function()` (equivalent to a `geom_` layer) to graph mathematical and statistical functions. All that is needed is a `data.frame` of a range of `x` values to act as the source for `data`, and set `x` equal to those values for `aes`thetics. 

```{r, fig.retina=3}
library(tidyverse)
# x values are integers 1 through 10
ggplot(data = data.frame(x = 1:10))+
  aes(x = x)
```

Then we add the `stat_function`, where `fun =` is the most important argument where you define the to function to graph as your function created above, for example, our `my_function`. 

```{r, fig.retina=3}
ggplot(data = data.frame(x = 1:10))+
  aes(x = x)+
  stat_function(fun = my_function) 
```

You can also adjust things like size, color, and line type.

```{r, fig.retina=3}
ggplot(data = data.frame(x = 1:10))+
  aes(x = x)+
  stat_function(fun = my_function, color = "blue", size = 2, linetype = "dashed") 
```

## Bultin Statistical Functions

There are some standard statistical distributions built into R. They require a combination of a specific prefix and a distribution. 

**Prefixes**:

| Action/Type | Prefix |
|--------|--------|
| random draw | `r` |
| density (pdf) | `d` |  
| cumulative density (cdf)    | `p`    |
| quantile (inverse cdf) | `q` |

**Distributions**:

| Distribution | Name in R |
|--------|--------|
| Normal | `norm` |
| Uniform | `unif` |  
| Student's t    | `t`    |
| Binomial    | `binom`    |
| Negative binomial    | `nbinom`    |
| Hypergeometric    | `hyper`    |
| Weibull    | `weibull`    |
| Beta    | `beta`    |
| Gamma    | `gamma`    |

Thus, what you want is a combination of the prefix and the distribution.

## Some common examples:

1. Take random draws from a normal distribution: 

```{r}
rnorm(n = 10, # take 10 draws from a normal distribution with:
      mean = 2, # mean of 2
      sd = 1) # sd of 1
```

2. Get probability of a random variable being less than or equal to a value (cdf) from a normal distribution: 

```{r}
# find probability of area to the LEFT of a number on pdf (note this = cdf of that number!)
pnorm(q = 80, # number is 80 from a distribution where: 
      mean = 200, # mean is 100
      sd = 100, # sd is 100
      lower.tail = TRUE) # looking to the LEFT in lower tail
```

3. Find the value of a distribution that is a specified percentile. 

```{r}
# find the 38th percentile value
qnorm(p = 0.38, # 38th percentile from a distribution where:
      mean = 200, # mean is 200
      sd = 100) # sd is 100
```

## Graphing Statistical Functions

You can also graph these commonly used statistical functions by setting `fun = ` the named functions in your `stat_function()` layer. If you want to specify the mean and standard deviation, use `args = list()` to include the required arguments from the named function above (e.g. `dnorm` needs `mean` and `sd`). 

```{r, fig.retina=3}
ggplot(data = data.frame(x = -400:600))+
  aes(x = x)+
  stat_function(fun = dnorm, args = list(mean = 200, sd = 200), color = "blue", size = 2, linetype = "dashed") 
```

If you don't include this, it will graph the standard distribution:

```{r, fig.retina=3}
ggplot(data = data.frame(x = -4:4))+
  aes(x = x)+
  stat_function(fun = dnorm, color = "blue", size = 2, linetype = "dashed") 
```

To add shading under a distribution, simply add a duplicate of the `stat_function()` layer, but add `geom="area"` to indicate the area beneath the function should be filled, and you can limit the domain of the `fill` with `xlim=c(start,end)`, where `start` and `end` are the x-values for the endpoints of the `fill.`

```{r, fig.retina=3}
# graph normal distribution and shade area between -2 and 2
ggplot(data = data.frame(x = -4:4))+
  aes(x = x)+
  stat_function(fun = dnorm, color = "blue", size = 2, linetype = "dashed")+
  stat_function(fun = dnorm, xlim = c(-2,2), geom = "area", fill = "green", alpha=0.5)
```

Hence, here is one graph from my slides:

```{r, fig.retina=3}

ggplot(data = tibble(x=35:115))+
  aes(x = x)+
  stat_function(fun = dnorm, args = list(mean = 75, sd = 10), size = 2, color="blue")+
  stat_function(fun = dnorm, args = list(mean = 75, sd = 10), geom = "area", xlim = c(65,85), fill="blue", alpha=0.5)+
  labs(x = "X",
       y = "Probability")+
  scale_x_continuous(breaks = seq(35,115,5))+
  theme_classic(base_family = "Fira Sans Condensed",
           base_size=20)
```