---
title: "3.6 — Regression with Categorical Data — Class Notes"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

*Tuesday, October 27, 2020*

# Overview

Today we look at how to use data that is categorical (i.e. variables that indicate an observation's membership in a particular group or category). We introduce them into regression models as **dummy variables** that can equal 0 or 1: where 1 indicates membership in a category, and 0 indicates non-membership.

We also look at what happens when categorical variables have more than two values: for regression, we introduce a dummy variable for each possible category - but be sure to leave out one reference category to avoid the dummy variable trap.

# Readings

Please see [today’s suggested readings](/reading/3.6-reading).

# Slides

- [<i class="fas fa-chalkboard-teacher"></i> View Lecture Slides](/slides/3.6-slides.html)
- [<i class="fas fa-file-pdf"></i> Download as PDF](/slides/3.6-slides.pdf)

# Assignments

[Problem Set 4](/assignment/04-problem-set) answers are posted on that page.

# Live Class Session on Zoom

The live class <i class="fas fa-video"></i> Zoom meeting link can be found on Blackboard (see `LIVE ZOOM MEETINGS` on the left navigation menu), starting at 11:30 AM.

If you are unable to join today's live session, or if you want to review, you can find the recording stored on Blackboard via Panopto (see `Class Recordings` on the left navigation menu).

# Appendix: T-Test for Difference in Group Means

Often we want to compare the means between two groups, and see if the difference is statistically significant. As an example, **is there a statistically significant difference in average hourly earnings between men and women**? Let:

- $\mu_W$: mean hourly earnings for female college graduates
- $\mu_M$: mean hourly earnings for male college graduates

We want to run a hypothesis test for the difference $(d)$ in these two population means: 
$$\mu_M-\mu_W=d_0$$

Our null hypothesis is that there is *no* statistically significant difference. Let's also have a two-sided alternative hypothesis, simply that there *is* a difference (positive or negative). 

- $H_0: d=0$
- $H_1: d \neq 0$

Note a logical one-sided alternative would be $H_2: d > 0$, i.e. men earn more than women

### The Sampling Distribution of $d$

The *true* population means $\mu_M, \mu_W$ are unknown, we must estimate them from *samples* of men and women. Let:
- $\bar{Y}_M$ the average earnings of a sample of $n_M$ men  
- $\bar{Y}_W$ the average earnings of a sample of $n_W$ women 

We then estimate $(\mu_M-\mu_W)$ with the sample $(\bar{Y}_M-\bar{Y}_W)$. 

We would then run a **t-test** and calculate the **test-statistic** for the difference in means. The formula for the test statistic is:

$$t = \frac{(\bar{Y_M}-\bar{Y_W})-d_0}{\sqrt{\frac{s_M^2}{n_M}+\frac{s_W^2}{n_W}}}$$

We then compare $t$ against the critical value $t^*$, or calculate the $p$-value $P(T>t)$ as usual to determine if we have sufficient evidence to reject $H_0$

```{r}
library(tidyverse)
library(wooldridge)
# Our data comes from wage1 in the wooldridge package

wages<-wooldridge::wage1

# look at average wage for men

wages %>%
  filter(female==0) %>%
  summarize(average = mean(wage),
            sd = sd(wage))

# look at average wage for women

wages %>%
  filter(female==1) %>%
  summarize(average = mean(wage),
            sd = sd(wage))

```

So our data is telling us that male and female average hourly earnings are distributed as such:

$$\begin{align*}
\bar{Y}_M &\sim N(7.10,4.16)\\
\bar{Y}_W &\sim N(4.59,2.53)\\
\end{align*}$$

We can plot this to see visually. There is a lot of overlap in the two distributions, but the male average is higher than the female average, and there is also a lot more variation in males than females, noticeably the male distribution skews further to the right. 

```{r}
wages$female<-as.factor(wages$female)

library("ggplot2")
ggplot(data=wages,aes(x=wage,fill=female))+
  geom_density(alpha=0.5)+
  scale_x_continuous(seq(0,25,5),name="Wage",labels=scales::dollar)+
  theme_light()
```

Knowing the distributions of male and female average hourly earnings, we can estimate the **sampling distribution of the difference in group eans** between men and women as: 

The mean: 
$$\begin{align*}
\bar{d}&=\bar{Y}_M-\bar{Y}_W\\
\bar{d}&=7.10-4.59\\
\bar{d}&=2.51\\
\end{align*}$$

The standard error of the mean:
$$\begin{align*}
SE(\bar{d})&=\sqrt{\frac{s_M^2}{n_M}+\frac{s_W^2}{n_W}}\\
&=\sqrt{\frac{4.16^2}{274}+\frac{2.33^2}{252}}\\
& \approx 0.29\\
\end{align*}$$

So the sampling distribution of the difference in group means is distributed:
$$\bar{d} \sim N(2.51,0.29)$$

```{r}
ggplot(data.frame(x=0:6),aes(x=x))+
  stat_function(fun=dnorm, args=list(mean=2.51, sd=0.29), color="purple")+
  ylab("Density")+
  scale_x_continuous(seq(0,6,1),name="Wage Difference",labels=scales::dollar)+
  theme_light()
```


Now we the **$t$-test** like any other: 

$$\begin{align*}
t&=\frac{\text{estimate}-\text{null hypothesis}}{\text{standard error of the estimate}}\\
&=\frac{d-0}{SE(d)}\\
&=\frac{2.51-0}{0.29}\\
&=8.66\\
\end{align*}$$

This is statistically significant. The $p$-value, $P(t>8.66)=$ is 0.000000000000000000410, or basically, 0. 


```{r}
pt(8.66,456.33, lower.tail=FALSE)
```

### The $t$-test in `R`

```{r}
t.test(wage~female, data=wages, var.equal=FALSE)
```

```{r}
reg<-lm(wage~female, data=wages)
summary(reg)
```