---
title: "Problem Set 2"
author: "YOUR NAME HERE" # put your name here! 
date: "ECON 480 — Fall 2020"
output: html_document # change to pdf if you'd like, this will make a webpage (you can email it and open in any browser)
---

<!--CLICK "KNIT" ABOVE TO RENDER TO HTML, PDF, OR WORD OUTPUT
In fact, try knitting right away and see what this produces!
-->

*Due by 11:59 PM Sunday September 13, 2020*

# Theory and Concepts

## Question 1
**In your own words, explain the difference between endogeneity and exogeneity.**

<!--WRITE YOUR ANSWERS BELOW -->

## Question 2

### Part A 

**In your own words, explain what (sample) standard deviation _means_.**

<!--WRITE YOUR ANSWERS BELOW -->

### Part B 

**In your own words, explain how (sample) standard deviation _is calculated._ You may also write the formula, but it is not necessary.**

<!--WRITE YOUR ANSWERS BELOW -->


# Problems

**For the remaining questions, you may use `R` to *verify*, but please calculate all sample statistics by hand and show all work.**

## Question 3

**Suppose you have a very small class of four students that all take a quiz. Their scores are reported as follows:**

$$\{83, 92, 72, 81\}$$

## Part A 
**Calculate the median.**

<!--WRITE YOUR ANSWERS BELOW -->

## Part B 
**Calculate the sample mean, $\bar{x}$**

<!--WRITE YOUR ANSWERS BELOW -->

## Part C 
**Calculate the sample standard deviation, $s$**

<!--WRITE YOUR ANSWERS BELOW -->

## Part D
**Make or sketch a rough histogram of this data, with the size of each bin being 10 (i.e. 70's, 80's, 90's, 100's). You can draw this by hand or use `R`.**^[If you are using `ggplot`, you want to use `+geom_histogram(breaks=seq(start,end,by))` and add `+scale_x_continuous(breaks=seq(start,end,by))`. For each, it creates bins in the histogram, and ticks on the x axis by creating a `seq`uence starting at `start` (a number), ending at `end` (number), `by` a certain interval (i.e. by `10`s.).] **Is this distribution roughly symmetric or skewed? What would we expect about the mean and the median?**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 5-c}
# write your code here
```


## Part E

**Suppose instead the person who got the 72 did not show up that day to class, and got a 0 instead. Recalculate the mean and median. What happened and why?**

<!--WRITE YOUR ANSWERS BELOW -->


## Question 4

Suppose the probabilities of a visitor to Amazon’s website buying 0, 1, or 2 books are 0.2, 0.4, and 0.4 respectively. 

### Part A
**Calculate the _expected number_ of books a visitor will purchase.**

<!--WRITE YOUR ANSWERS BELOW -->

### Part B
**Calculate the _standard deviation_ of book purchases.**

<!--WRITE YOUR ANSWERS BELOW -->

### Part C 
**BONUS: try doing this in `R` by making an initial dataframe of the data, and then making new columns to the "table" like we did in class.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 4-c}
# write your code here
```

## Question 5

Scores on the SAT (out of 1600) are approximately normally distributed with a mean of 500 and standard deviation of 100. 

### Part A
**What is the probability of getting a score between a 400 and a 600?**

<!--WRITE YOUR ANSWERS BELOW -->

### Part B
**What is the probability of getting a score between a 300 and a 700?**

<!--WRITE YOUR ANSWERS BELOW -->

### Part C
**What is the probability of getting _at least_ a 700?**

<!--WRITE YOUR ANSWERS BELOW -->

### Part D
**What is the probability of getting _at most_ a 700?**

<!--WRITE YOUR ANSWERS BELOW -->

### Part E
**What is the probability of getting _exactly_ a 500?**

<!--WRITE YOUR ANSWERS BELOW -->

## Question 6

**Redo problem 5 by using the `pnorm()` command in `R`.**^[Hint: This function has four arguments: 1. the value of the random variable, 2. the mean of the distribution, 3. the sd of the distribution, and 4. `lower.tail` `TRUE` or `FALSE`.]

<!--WRITE YOUR ANSWERS BELOW -->

```{r 6}
# write your code here
```